Millions of Type Errors

### 2023-08-31 programming23/paper.tex
checkpoint, data overview

\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\FILL}{\textbf{FILL}}
\newcommand{\dotscale}[1]{\scalebox{0.72}{#1}}
\newcommand{\wideas}[2]{\makebox[\widthof{#2}][l]{#1}}
\newcommand{\twoline}[2]{\parbox[s]{1.4cm}{\flushleft#1\newline#2}}
\newcommand{\chkYes}{\dotscale{\CIRCLE}}
\newcommand{\chkMaybe}{\wideas{\dotscale{\Circle}}{\chkYes}}
\newcommand{\chkNo}{\wideas{}{\chkYes}}
\newcommand{\pct}[1]{\SI{#1}{\percent}}
\newcommand{\modefont}[1]{\texttt{#1}}
\newcommand{\mnocheck}{\modefont{nocheck}}
\newcommand{\mnonstrict}{\modefont{nonstrict}}
\newcommand{\mstrict}{\modefont{strict}}

@@

\begin{figure}[t]
  \includegraphics[width=\columnwidth]{img/row-distribution.pdf}
  \caption{Telemetry records per hour. Each tick on the $x$-axis marks the start of a new day in California. Shaded regions correspond to weekends.}
  \label{f:records-per-hour}
\end{figure}

@@

\begin{figure}[t]
  \begin{tabular}{rl}
   5,001 & total logs \\
         & 4,471 \mnocheck{} + 506 \mnonstrict{} + 24 \mstrict{} \\
         & 1,628 from module switches \\
   182,247 & total forced strict type errors \\
         & {83,005 in module, 5,113 in edit regions} \\
    1,521 & total type errors \\
         & {1,035 in module, 110 non-syntax + 47 syntax in edit regions}
    \\[2ex]

    1,146 & sessions \\
    & \begin{tabular}{l}
      1,143 single mode = 1046 \mnocheck{} + 94 \mnonstrict{} + 3 \mstrict{} \\
        detected 2 multi-mode projects, 2 mode upgrades, 2 mode downgrades
      \end{tabular}
  \end{tabular}

  %% TODO these need to be plots, it's too much reading for the quick takeaway
  %% put exactly this data into bars!
  \begin{minipage}[t]{0.5\textwidth}
    \begin{tabular}{rr}
      Session length & \# Sessions \\\midrule
      1 &  450 \\
      2 &  199 \\
      3 &  129 \\
      4 &  73 \\
      5 &  60 \\
      6-10 & 150 \\
      11-20 & 67 \\
      21-40 & 11 \\
      >40 & 7
      %% max 393
    \end{tabular}
  \end{minipage}\begin{minipage}[t]{0.5\textwidth}
    \begin{tabular}{rr}
      Time delta & \# Records \\\midrule
      0-5ms & 284 \\
      5-1000ms & 248 \\
      1-2s & 153 \\
      2-5s & 232 \\
      5-10s & 207 \\
      10-30s & 524 \\
      30-60s & 392 \\
      1-2m & 394 \\
      2-5m & 559 \\
      5-10m & 360 \\
      >10-60m & 502
    \end{tabular}
  \end{minipage}

@@

\subsection{Dead Ends: Too Complex}

The typechecker has two internal limits regarding the problems it will attempt
to solving.
Hitting a limit triggers either a \code{CodeTooComplex} or
\code{UnificationTooComplex} error.
A casual user should never see these errors.

Based on the data, these errors are indeed extremely rare.
Out of 1.5M telemetry records, only 11 report any \code{TooComplex} errors.
These records came from three sessions (possibly by a single user), and each
record counted at most 4 \code{CodeTooComplex} errors and zero
\code{UnificationTooComplex} errors.
There is no evidence to suggest that the limits are too strict or that the
typechecker should improve.

>>>>

\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\FILL}{\textbf{FILL}}
\newcommand{\dotscale}[1]{\scalebox{0.72}{#1}}
\newcommand{\wideas}[2]{\makebox[\widthof{#2}][l]{#1}}
\newcommand{\twoline}[2]{\parbox[s]{1.4cm}{\flushleft#1\newline#2}}
\newcommand{\chkYes}{\dotscale{\CIRCLE}}
\newcommand{\chkMaybe}{\wideas{\dotscale{\Circle}}{\chkYes}}
\newcommand{\chkNo}{\wideas{}{\chkYes}}
\newcommand{\pct}[1]{\SI{#1}{\percent}}
\newcommand{\modefont}[1]{\texttt{#1}}
\newcommand{\mnocheck}{\modefont{nocheck}}
\newcommand{\mnonstrict}{\modefont{nonstrict}}
\newcommand{\mstrict}{\modefont{strict}}
\newcommand{\zerowidth}[1]{\makebox[0pt][l]{#1}}

@@

\begin{figure}[t]
  %% TODO bigger text
  %% code/row-distribution.rkt
  \includegraphics[width=\columnwidth]{img/row-distribution.pdf}
  \caption{Telemetry records per hour. Each tick on the $x$-axis marks the start of a new day in California. Shaded regions correspond to weekends.}
  \label{f:records-per-hour}
\end{figure}

@@

"89.1417"
"10.4259"
"0.4323"

  %% TODO subfigure??
  \begin{tabular}{ll}
    \begin{tabular}{r@{~~}l@{~~}r}
       1,341,348 & \mnocheck{}          & [\pct{89.14}] \\
         156,883 & \mnonstrict{}        & [\pct{10.43}] \\
           6,505 & \mstrict{}           & [\pct{ 0.43}] \\\midrule
       1,504,736 & \zerowidth{Total telemetry records}
    \end{tabular}
    \begin{tabular}{r@{~~}l@{~~}r}
         508,572 & due to module switch & [\pct{33.80}] \\
         996,164 & due to keystrokes    & [\pct{66.20}]
    \end{tabular}
  \end{tabular}

 72,235,735 total forced strict type errors, 37,027,281 in module, 1,111,178 in edit regions

 595,137 total type errors, 289,698 in module, 14,917 non-stx + 16,007 syntax in edit regions

 347,598 sessions
  346,956 single mode = 313,509 nocheck + 32,902 nonstrict + 545 strict
  512 multi mode projects
  341 mode upgrades
  320 mode downgrades

%  \begin{tabular}{rl}
%   182,247 & total forced-strict type errors \\
%         & {83,005 in module, 5,113 in edit regions} \\
%    1,521 & total type errors \\
%         & {1,035 in module, 110 non-syntax + 47 syntax in edit regions}
%    \\[2ex]
%
%    1,146 & sessions \\
%    & \begin{tabular}{l}
%      1,143 single mode = 1046 \mnocheck{} + 94 \mnonstrict{} + 3 \mstrict{} \\
%        detected 2 multi-mode projects, 2 mode upgrades, 2 mode downgrades
%      \end{tabular}
%  \end{tabular}

%  %% TODO these need to be plots, it's too much reading for the quick takeaway
%  %% put exactly this data into bars!
%  \begin{minipage}[t]{0.5\textwidth}
%    \begin{tabular}{rr}
%      Session length & \# Sessions \\\midrule
%      1 &  450 \\
%      2 &  199 \\
%      3 &  129 \\
%      4 &  73 \\
%      5 &  60 \\
%      6-10 & 150 \\
%      11-20 & 67 \\
%      21-40 & 11 \\
%      >40 & 7
%      %% max 393
%    \end{tabular}
%  \end{minipage}\begin{minipage}[t]{0.5\textwidth}
%    \begin{tabular}{rr}
%      Time delta & \# Records \\\midrule
%      0-5ms & 284 \\
%      5-1000ms & 248 \\
%      1-2s & 153 \\
%      2-5s & 232 \\
%      5-10s & 207 \\
%      10-30s & 524 \\
%      30-60s & 392 \\
%      1-2m & 394 \\
%      2-5m & 559 \\
%      5-10m & 360 \\
%      >10-60m & 502
%    \end{tabular}
%  \end{minipage}

@@

\subsection{Dead Ends: Too Complex}
%% data = out/ctc-info.txt

To deal with pathologies such as the worst-case time for ML-style type
inference~\cite{m-popl-1990,ktu-caap-1990}, the typechecker
has two internal limits to restrict the problems it will attempt to solve.
Hitting a limit triggers either a \code{CodeTooComplex} or
\code{UnificationTooComplex} error.
A casual user should never see these errors.

Fortunately, the data rarely contains too-complex errors.
There are zero \code{UnificationTooComplex} errors and only a
handful of \code{CodeTooComplex} errors.
%% 26 total
It is entirely possible that the \code{CodeTooComplex} errors came from
a single user, as they appear in eleven records (out of 1.5 million)
spread across three sessions.
At any rate, the overall number is low enough to suggest that the internal
limits are not posing barriers to typechecker adoption.
